---
title: "R Notebook"
output: html_notebook
---

# Some notes on the data
**survived:**
0 = No
1 = Yes

**pclass**: A proxy for socio-economic status (SES)
1st = Upper
2nd = Middle
3rd = Lower

**age**: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

**sibsp**: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

**parch:** The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.

**embarked:** what is the port where the person embarked from
C = Cherbourg, Q = Queenstown, S = Southampton

```{r libraries}
library(ggplot2)
library(dplyr)
library(corrplot)
library(fastDummies)
library(ROCR)
```

```{r data_loads}
# reading the data of the competition
train <- read.csv('all/train.csv')
test <- read.csv('all/test.csv')

names(train)
train$is_train <- TRUE
test$is_train <- FALSE
test$Survived <- NA

# let's make sure that we do the same transformations to both training and test ...
train <- bind_rows(train, test)
```

```{r basic_exploration}
# types of data
str(train)

# size of the dataset
dim(train)
# 891 rows and 12 variables

# amount of NA's per column
colSums(is.na(train))[colSums(is.na(train)) > 0]
# only Age has NA's

# how balanced is the dataset ...
round(table(train$Survived) / dim(train)[1] * 100, 1)
```

Some points about the data:
- 891 Rows, 12 Variables
- Only Age has NA's ... and it's 177 our of the 891 rows of data!
- 61.6% doesn't survive, 38.4% survives

```{r dataTransformations}

# Family properties  ----
# CHILDs ====
# we separate children from adults, we use the age as a variable. And stablish that <16 is a child.
# we verify this assumption by making sure that children labeled don't have more than 3 parents...
# we also know that people with n_parents = 0 were travelling with a nanny; therefore being children. So we also use that rule

# Verifying child assumption
train %>%
  mutate(
    child = (Age < 16) * 1 | (Parch == 0 & Age < 16)
  ) %>%
  group_by(child) %>%
  summarise(
    n_parents = paste(unique(Parch), collapse = ", ")
  )
# people without a label (NA) just like children, have a maximum of 2 parents... could these people have possibly been children?

# let's analyze the distribution of Age
train %>%
  filter(is_train == TRUE) %>%
  group_by(Pclass, Sex) %>%
  summarise(
    median_Age = median(Age, na.rm = T)
  ) -> ageMedians

# We can see that different classes and sexes have different medians of ages...
# we shall replace NA's in accordance to sexes and classes, so that it reflects better the age distribution of their characteristics
(train %>%
  left_join(
    ageMedians
  ) %>%
  mutate(
    inputAge = ifelse(is.na(Age), median_Age, Age)
  ))$inputAge -> inputAge

# Name information ====
# the name tag has information about marital status, job (Dr., Col, etc.) and other features (Master?)

# luckily, all of the titles of either the job or marital status have a . at the end. We can easily extract them
getDotName <- function(x){
  strsplit(x, " ")
  x[grep("\\.", x)]
}

nameList <- strsplit(as.character(train$Name), " ")
nameList <- lapply(nameList, function(x) x[grep("\\.", x)][1])
nameList <- unlist(nameList)
rev(sort(table(nameList)))

# several new variables can be created...
# spouse, military, master, Dr., Rev.

# Cabin info ====
rev(sort(table(train$Cabin)))
# 687 persons have no cabin register... the 200 less have cabins all over the place. We will simplify and create two variables:
# the letter for the section
# the number for the cabin location
# some passengers have multiple cabin information, to simplify we just grab the first letter

# getting cabin numbers ####
getCabinNumber <- function(x){
  strsplit(x, " ")
}

cabinNumbers <- lapply(as.character(train$Cabin), getCabinNumber)
cabinNumbers <- sapply(cabinNumbers, function(x) x[[1]][1])
cabinNumbers <- gsub("[^0-9]", "", cabinNumbers)
cabinNumbers[cabinNumbers == ''] <- NA
cabinNumbers <- as.numeric(cabinNumbers)
#medianCabin <- median(as.numeric(cabinNumbers[!is.na(cabinNumbers)]), na.rm = T)
#cabinNumbers[is.na(cabinNumbers)] <- medianCabin

# FINAL TRANSFORMATION ----
# all of the transformations
train %>% mutate(
  Age = inputAge,
  child = ((Age < 16) | (Parch == 0 & Age < 16)) * 1,
  spouse = (nameList %in% c("Mrs.", "Mlle.", "Lady.")) * 1,
  single_woman = (nameList %in% c("Miss.", "Mmme.")) * 1,
  master = (nameList == "Master.") * 1,
  job_title = nameList %in% c("Dr.", "Rev.", "Major.", "Col.", "Capt.", "L.") * 1,
  noble = nameList %in% c("Sir.", "Lady.", "Don.", "Countess.", "Jonkheer") * 1,
  mister = (nameList %in% 'Mr.') * 1,
  cabin_letter = ifelse(substr(Cabin, 1, 1) == '', NA, substr(Cabin, 1, 1)),
  cabin_numbers = cabinNumbers,
  male = (Sex == "male") * 1,
  Sex = NULL,
  cabin_letter = as.factor(cabin_letter)
) -> train2

# we're going to get rid of columns that don't provide value
train2$PassengerId <- NULL
train2$Name <- NULL
train2$Cabin <- NULL
train2$Ticket <- NULL

# Let's transform the dataset into a dummy dataframe
dummy_df <- dummy_cols(Filter(is.factor, train2), remove_first_dummy = T)

train2 <- cbind(train2, dummy_df[,-c(1, 2)])

train2$Embarked <- NULL
train2$cabin_letter <- NULL
```

We're doing some data exploration now, to understand the relationship between the different variables

```{r data_exploration}
# get only the numeric variables
corrplot(cor(Filter(is.numeric, train2)))

numeric_df <- Filter(is.numeric, train2)
dim(numeric_df)

# we can see that we have a couple of metrics that are correlated... we will get rid of them.
train2$mister <- NULL
```

Interesting findings:
1- Cabin A seems to be correlated with noble people and neg corr with class. Meaning probably most higher class person were staying in this cabin number
2- Master and child seem to be very correlated, (Just found out this was a way of calling kids, whoops!)
3- Mister. and male sex are correlated (Which is obvious)
4- Most cabin information seems to be correlated similarly, implying that those who had cabins registered probably share a lot of characteristics. We will most likely elliminate or simplify this metric after checking it's performance within a model

```{r dist plots}
par(mfrow = c(2, 2))
for(i in 1:dim(numeric_df)[2]){
  if(length(unique(numeric_df[,i])) < 10){
   barplot(table(unlist(train2[,i])), main = names(numeric_df)[i]) 
  }else{
   hist(unlist(train2[,i]), main = names(numeric_df)[i]) 
  }
}
```

Model trainings
```{r model_training}
# let's separate in train, test and validation
randomDraw <- runif(dim(train2)[1]) < 0.6
train2$has_cabin <- rowSums(train2[,17:22])
train2[,17:22] <- NULL
train2$cabin_numbers <- NULL
train2$cabin_letter_T <- NULL

# scaling the data!!!
train2 %>% select(Age, Fare) %>% scale() -> scaled_data

scaled_data <- matrix(scaled_data, ncol = 2)
scaled_data <-  data.frame(
  'age_scaled' = scaled_data[,1],
  'fare_scaled' = scaled_data[,2]
)
train2 <- cbind(train2, scaled_data)
train2$Age <- NULL
train2$Fare <- NULL

train2 %>% filter(is_train == F) -> validation
train2 %>% filter(is_train == T & randomDraw == T) -> train
train2 %>% filter(is_train == T & randomDraw == F) -> test

# let's try a  logistic model, random forest and xgboost
mylogit <- glm(Survived ~ ., data = train, family = "binomial")
steps_mylogit <- step(mylogit)

# these are the steps I want to keep
#mylogit <- glm(formula = Survived ~ Pclass + Age + SibSp + Parch + Fare + 
#    single_woman + master + male + has_cabin, family = "binomial", 
#    data = train)

mylogit2 <- glm(formula = Survived ~ Pclass + Age + SibSp + single_woman + 
    master + male + has_cabin, family = "binomial", data = train)

predictions <- (predict(mylogit, test,  type = 'response') > 0.5) * 1
predictions2 <- (predict(mylogit2, test,  type = 'response') > 0.5) * 1

accuracy <- 1 - mean(predictions != test$Survived)
accuracy2 <- 1 - mean(predictions2 != test$Survived)
```

```{r performance_logit_regression}
p <- predict(mylogit2, test, type="response")
pr <- prediction(p, test$Survived)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

